---
title: "S4E5"
author: "UVM SAC"
date: "2023-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center")
```

## S4E5 - Basketball Data Visualization

In this weeks workshop, we're going to be working with R using an R markdown file, which combines markdown(formatted text like this one) and code chunks. Hopefully everyone can get their own R environment set up so that anyone can experiment with the data and make their own charts.

The package we're working with today is called hoopR
```{r package_setup}
# installing the package
#install.packages("hoopR")
#install.packages("sf")
# initializing the package for this R session
library(hoopR)
library(tidyverse)
library(sf)
```

hoopR is an open source package that contains up to date NBA and NCAA play-by-play data

```{r loading_data}
# load in the ncaa play-by-play
# data from 2006-2023

# its a large dataset, so it may take a while to run
ncaa_pbp <- load_mbb_pbp(seasons = 2023)

shot_types = c("DunkShot", "JumpShot", "LayUpShot", "TipShot")
```

```{r}
# plotting shot chart for one team in one game

ncaa_pbp |> 
  filter(game_id == 401522202,#2023 national championship
         type_text %in% shot_types) |> 
  ggplot(
    mapping = aes(
      x = coordinate_x,
      y = coordinate_y
    )
  ) + 
  geom_point(
    mapping = aes(
      color = factor(score_value),
      shape = scoring_play
    ),
    size = 2
  ) + 
  scale_shape_manual(
    values = c(4, 16)
  ) + coord_fixed()
# home team is on the right, away on the left
```


Lets try adding the lines on the court to better make sense of the data

Below will be a lot of code that uses spacial objects to create to-scale lines for an NCAA court. The final result will be a theme that we can add to our shot chart graph that draws in the lines with only one line of code.

```{r}
library(sf) # Work with polygons

# All lengths will be in feet
line_thick = 0.15
width = 50
height = 94 / 2
key_height = 19
key_width = 12
key_radius = 6
backboard_width = 6
# https://www.fiba.basketball/documents/BasketballEquipment.pdf
backboard_thick = 0.3
backboard_offset = 4
hoop_radius = 0.75
hoop_center_y = 5.25
rim_thick = 0.06
neck_length = hoop_center_y - (backboard_offset + hoop_radius + rim_thick)
three_point_radius = 22 + .1458
three_point_side_offset = 3 + .33 + .01
three_point_side_height = sqrt(
  three_point_radius^2 - (three_point_side_offset - width/2)^2
  ) + hoop_center_y
restricted_area_radius = 4

# Draw a rectangle that defines the half-court interior
half_court_int <- rbind(
  c(0, 0),
  c(0, height),
  c(width, height),
  c(width, 0),
  c(0,0)
  )

# Draw a rectangle that defines the half-court exterior
half_court_ext <- rbind(
  c(0-line_thick, 0-line_thick),
  c(0-line_thick, height + line_thick),
  c(width + line_thick, height + line_thick),
  c(width + line_thick, 0-line_thick),
  c(0-line_thick, 0-line_thick)
  )

# Define a sfg polygon object in sf by subtracting interior from exterior
half_court <- st_polygon(list(half_court_ext, half_court_int))


# Draw a rectangle for the key
key_int <- rbind(
  c(width/2 - key_width/2 + line_thick, 0),
  c(width/2 - key_width/2 + line_thick, key_height - line_thick),
  c(width/2 + key_width/2 - line_thick, key_height - line_thick),
  c(width/2 + key_width/2 - line_thick, 0),
  c(width/2 - key_width/2 + line_thick, 0)
)

key_ext <- rbind(
  c(width/2 - key_width/2, 0),
  c(width/2 - key_width/2, key_height),
  c(width/2 + key_width/2, key_height),
  c(width/2 + key_width/2, 0),
  c(width/2 - key_width/2, 0)
)

key <- st_polygon(list(key_ext, key_int))

# Draw a rectangle for the backboard
backboard_points <- rbind(
  c(width/2 - backboard_width/2, backboard_offset - backboard_thick),
  c(width/2 - backboard_width/2, backboard_offset),
  c(width/2 + backboard_width/2, backboard_offset),
  c(width/2 + backboard_width/2, backboard_offset - backboard_thick),
  c(width/2 - backboard_width/2, backboard_offset - backboard_thick)
)

backboard <- st_polygon(list(backboard_points))

# Neck
neck_points <- rbind(
  c(width/2 - line_thick/2, backboard_offset),
  c(width/2 - line_thick/2, backboard_offset + neck_length),
  c(width/2 + line_thick/2, backboard_offset + neck_length),
  c(width/2 + line_thick/2, backboard_offset),
  c(width/2 - line_thick/2, backboard_offset)
)

neck <- st_polygon(list(neck_points))

# Define a point sfg object for the center of the hoop
hoop_center <- st_point(c(width/2, hoop_center_y))

# Interior of the rim
# Buffer the point by the radius of the hoop to create a circle
hoop_int <- hoop_center %>%
  st_buffer(dist = hoop_radius)

# Exterior of the rim
hoop_ext <- hoop_center %>%
  st_buffer(dist = hoop_radius + rim_thick)

# Subtract interior from exterior to get the rim
hoop <- st_polygon(list(
  # Only kepp the X, Y columns of the coordinates
  st_coordinates(hoop_ext)[ , 1:2],
  st_coordinates(hoop_int)[ , 1:2]
  ))

# Draw the half-circle at the top of the key
key_center <- st_point(c(width/2, key_height))

key_circle_int <- st_crop(
  st_sfc(st_buffer(key_center, dist = key_radius - line_thick)),
  # Only keep the part of the circle above the top of the key
  xmin = 0, ymin = key_height, xmax = width, ymax = height
  )

key_circle_ext <- st_crop(
  st_sfc(st_buffer(key_center, dist = key_radius)),
  xmin = 0, ymin = key_height, xmax = width, ymax = height
)

key_circle <- st_polygon(list(
  st_coordinates(key_circle_ext)[ , 1:2],
  st_coordinates(key_circle_int)[ , 1:2]
  ))

# Draw the half-circle at the bottom of half-court
half_center <- st_point(c(width/2, height))

half_circle_int <- st_crop(
  st_sfc(st_buffer(half_center, dist = key_radius - line_thick)),
  # only keep the bottom half below the half-court line
  xmin = 0, ymin = 0, xmax = width, ymax = height
)

half_circle_ext <- st_crop(
  st_sfc(st_buffer(half_center, dist = key_radius)),
  xmin = 0, ymin = 0, xmax = width, ymax = height
)

half_circle <- st_polygon(list(
  st_coordinates(half_circle_ext)[ , 1:2],
  st_coordinates(half_circle_int)[ , 1:2]
))

# Define a point sfg object for the center of the hoop
three_center <- st_point(c(width/2, hoop_center_y))

# Buffer the point to create a circle & crop it at 2.99 meters
three_int <- st_crop(
  st_sfc(st_buffer(three_center, dist = three_point_radius - line_thick)),
  xmin = three_point_side_offset + line_thick, ymin = three_point_side_height,
  xmax = width - (three_point_side_offset + line_thick), ymax = height
)

# Get the number of rows of coordinates of the three_int object
n <- nrow(st_coordinates(three_int))

# Bind the straight line points to the arc
three_int <- rbind(
  c(three_point_side_offset + line_thick, 0),
  c(three_point_side_offset + line_thick, three_point_side_height),
  # Remove the last two rows and only keep the X,Y columns
  st_coordinates(three_int)[1:(n-2), 1:2],
  c(width - (three_point_side_offset + line_thick), three_point_side_height),
  c(width - (three_point_side_offset + line_thick), 0),
  c(three_point_side_offset + line_thick, 0)
)

# Do the same for the exterior
three_ext <- st_crop(
  st_sfc(st_buffer(three_center, dist = three_point_radius)),
  xmin = three_point_side_offset, ymin = three_point_side_height,
  xmax = width - three_point_side_offset, ymax = height
)

three_ext <- rbind(
  c(three_point_side_offset, 0),
  c(three_point_side_offset, three_point_side_height),
  st_coordinates(three_ext)[1:(n-2), 1:2],
  c(width - three_point_side_offset, three_point_side_height),
  c(width - three_point_side_offset, 0),
  c(three_point_side_offset, 0)
)

# Create a three-point line sfg polygon object
three_point_line <- st_polygon(list(three_int, three_ext))

# Restricted area
ra_center <- st_point(c(width/2, hoop_center_y))

ra_ext <- st_crop(
  st_sfc(st_buffer(ra_center, dist = restricted_area_radius + line_thick)),
  xmin = 0, ymin = hoop_center_y,
  xmax = width, ymax = height
)

n <- nrow(st_coordinates(ra_ext))

ra_ext <- tibble(
  x = st_coordinates(ra_ext)[1:(n-2), 1],
  y = st_coordinates(ra_ext)[1:(n-2), 2]
)


ra_ext <- rbind(
  c(width/2 - restricted_area_radius - line_thick, backboard_offset),
  c(width/2 - restricted_area_radius - line_thick, hoop_center_y),
  ra_ext,
  c(width/2 + restricted_area_radius + line_thick, hoop_center_y),
  c(width/2 + restricted_area_radius + line_thick, backboard_offset)
)

ra_int <- st_crop(
  st_sfc(st_buffer(ra_center, dist = restricted_area_radius)),
  xmin = 0, ymin = hoop_center_y,
  xmax = width, ymax = height
)

# Reverse the direction of the interior arc points
ra_int_flip <- tibble(
  x = st_coordinates(ra_int)[1:(n-2), 1],
  y = st_coordinates(ra_int)[1:(n-2), 2]
) %>%
  arrange(desc(x))

ra_int <- rbind(
  c(width/2 + restricted_area_radius, backboard_offset),
  c(width/2 + restricted_area_radius, hoop_center_y),
  ra_int_flip,
  c(width/2 - restricted_area_radius, hoop_center_y),
  c(width/2 - restricted_area_radius, backboard_offset),
  c(width/2 - restricted_area_radius - line_thick, backboard_offset)
)

# Bind all the points together
ra_points <- as.matrix(rbind(ra_ext, ra_int))

restricted_area <- st_polygon(list(ra_points))

# Create sf object with 9 features and 1 field
court_sf <- st_sf(
  description = c("half_court", "key", "hoop", "backboard",
                  "neck", "key_circle", "three_point_line",
                  "half_circle", "restricted_area"), 
  geom = c(st_geometry(half_court), st_geometry(key), st_geometry(hoop),
           st_geometry(backboard), st_geometry(neck), st_geometry(key_circle), 
           st_geometry(three_point_line), st_geometry(half_circle),
           st_geometry(restricted_area))
)

court_themes = list(
  light = list(
    court = '#ffffff',
    lines = '#000000',
    text = '#222222',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 0.3,
    hex_border_color = "#cccccc"
  ),
  dark = list(
    court = '#000004',
    lines = '#ffffff',
    text = '#f0f0f0',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 0,
    hex_border_color = "#000000"
  )
)

plot_court = function(court_theme = court_themes$light) {
    ggplot() + 
    geom_sf(data = court_sf,
            fill = court_theme$lines, col = court_theme$lines) +
    theme_void() +
    theme(
      text = element_text(color = court_theme$text),
      plot.background = element_rect(
        fill = court_theme$court, color = court_theme$court),
      panel.background = element_rect(
        fill = court_theme$court, color = court_theme$court),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      legend.background = element_rect(
        fill = court_theme$court, color = court_theme$court),
      legend.margin = margin(-1, 0, 0, 0, unit = "lines"),
      legend.position = "bottom",
      legend.key = element_blank(),
      legend.text = element_text(size = rel(1.0))
    )
}
```

Implementing the court lines:
We will have to make some changes to the graph to align with the theme. For the court lines, (0,0) is in the bottom left corner. In our previous graph (0, 0) is mid court. Also, with the court theme we only want to look at one team at a time. Also, for the theme, the x and y axes are going to be flipped.
```{r}
plot_court()
```

Adding shot charts to plot_court()
```{r}
ncaa_pbp |> 
  filter(
         team_id == 41, # UConn
         type_text %in% shot_types) -> uconn_data

plot_court() +
  geom_point(
    data = uconn_data,
    mapping = aes(
      color = factor(score_value), 
      x = coordinate_y + 25, # adjusts for different coordinates
      y = -1 * abs(coordinate_x) + 47, # Uses correct coordinate system and all shots are on the same half
      shape = scoring_play
    ),
    size = 2
  ) + 
  scale_shape_manual(
    values = c(4, 16),
    labels = c("Miss", "Make")
  ) + 
  theme(legend.position = "right") + 
  labs(title = "2023 UConn Huskies Shot Chart",
       shape = NULL,
       color = NULL)


```

The data gets hard to read when you want to look at an entire season's worth of shots. To solve this we can try segmenting the court and look at shooting statistics for each zone.

```{r}
key_ext <- rbind(
  c(width/2 - key_width/2, 0),
  c(width/2 - key_width/2, key_height),
  c(width/2 + key_width/2, key_height),
  c(width/2 + key_width/2, 0),
  c(width/2 - key_width/2, 0)
)

three_ext <- st_crop(
  st_sfc(st_buffer(three_center, dist = three_point_radius)),
  xmin = three_point_side_offset, ymin = three_point_side_height,
  xmax = width - three_point_side_offset, ymax = height
)
n <- dim(st_coordinates(three_ext))[1]


# Left Corner 3
left_corner_three <- rbind(
  c(width - three_point_side_offset, 0),
  c(width - three_point_side_offset, three_point_side_height),
  c(width, three_point_side_height),
  c(width, 0),
  c(width - three_point_side_offset, 0)
)

left_corner_three <- st_polygon(list(left_corner_three))

# Right Corner 3
right_corner_three <- rbind(
  c(0, 0),
  c(0, three_point_side_height),
  c(three_point_side_offset, three_point_side_height),
  c(three_point_side_offset, 0),
  c(0, 0)
)

right_corner_three <- st_polygon(list(right_corner_three))

# Mid-Range Left
three_ext <- tibble(
  x = st_coordinates(three_ext)[1:(n-2), 1],
  y = st_coordinates(three_ext)[1:(n-2), 2]
)

mid_range_left <- rbind(
  c(three_point_side_offset, three_point_side_height),
  filter(three_ext, x <= (width/2 - key_width/2) + 1),
  c(width/2 - key_width/2, key_height),
  c(width/2 - key_width/2, three_point_side_height),
  c(three_point_side_offset, three_point_side_height)
)

mid_range_left <- st_polygon(list(as.matrix(mid_range_left)))


# Mid-Range Right

mid_range_right <- rbind(
  c(width - three_point_side_offset, three_point_side_height),
  arrange(filter(three_ext, x >= (width/2 + key_width/2) - 1), desc(x)),
  c(width/2 + key_width/2, key_height),
  c(width/2 + key_width/2, three_point_side_height),
  c(width - three_point_side_offset, three_point_side_height)
)

mid_range_right <- st_polygon(list(as.matrix(mid_range_right)))


# Mid-Range center

mid_range_center <- rbind(
  c(width/2 - key_width/2, key_height),
  filter(three_ext, x >= (width/2 - key_width/2 ), x <= (width/2 + key_width/2)),
  c(width/2 + key_width/2, key_height),
  c(width/2 - key_width/2, key_height)
)

mid_range_center <- st_polygon(list(as.matrix(mid_range_center)))


# Right Block
right_block <- rbind(
  c(width/2 + key_width/2, 0),
  c(width/2 + key_width/2, three_point_side_height),
  c(width - three_point_side_offset, three_point_side_height),
  c(width - three_point_side_offset, 0),
  c(width/2 + key_width/2, 0)
)

right_block <- st_polygon(list(as.matrix(right_block)))

# left Block
left_block <- rbind(
  c(width/2 - key_width/2, 0),
  c(width/2 - key_width/2, three_point_side_height),
  c(three_point_side_offset, three_point_side_height),
  c(three_point_side_offset, 0),
  c(width/2 - key_width/2, 0)
)

left_block <- st_polygon(list(as.matrix(left_block)))

# Restricted Area
ra_ext <- rbind(
  ra_ext,
  c(width/2 - restricted_area_radius - line_thick, backboard_offset)
)

restricted_area <- st_polygon(list(as.matrix(ra_ext)))

# In The Paint (Non-RA)
key_ext <- st_polygon(list(key_ext))

paint <- st_difference(key_ext, restricted_area)




deep_three_arc <- st_crop(
  st_sfc(st_buffer(three_center, dist = three_point_radius + 5)),
  xmin = 0, ymin = key_height - (key_height - three_point_side_height)/2,
  xmax = width, ymax = height
)

# Get the number of rows of coordinates of the three_int object
n <- nrow(st_coordinates(deep_three_arc))

# Bind the straight line points to the arc

#deep_three <- rbind(
#  c(0, key_height - (key_height - three_point_side_height)/2),
#  # Remove the last two rows and only keep the X,Y columns
#  st_coordinates(deep_three)[1:(n-2), 1:2],
#  c(width, key_height - (key_height - three_point_side_height)/2),
#  c(0, key_height - (key_height - three_point_side_height)/2)
#)

deep_three_arc <- tibble(
  x = st_coordinates(deep_three_arc)[1:(n-2), 1],
  y = st_coordinates(deep_three_arc)[1:(n-2), 2]
)



# regular three
three_left <- rbind(
  c(0, three_point_side_height),
  arrange(filter(deep_three_arc, x <= (width/2 - key_width/2)*.65), x),
  arrange(filter(three_ext, x <= (width/2 - key_width/2)*.75), desc(x)),
  c(three_point_side_offset, three_point_side_height),
  c(0, three_point_side_height)
)

three_left <- st_polygon(list(as.matrix(three_left)))

three_middle <- rbind(
  arrange(filter(deep_three_arc, x >= (width/2 - key_width/2)*.6, x <= (width/2 + key_width/2) + .4*(width - (width/2 + key_width/2))), x),
  arrange(filter(three_ext, x >= (width/2 - key_width/2)*.7, x <= (width/2 + key_width/2) + .3*(width - (width/2 + key_width/2))), desc(x)),
  arrange(filter(deep_three_arc, x >= (width/2 - key_width/2)*.6), x)[1,]
)

three_middle <- st_polygon(list(as.matrix(three_middle)))


three_right <- rbind(
  c(width, three_point_side_height),
  arrange(filter(deep_three_arc, x >= (width/2 + key_width/2) + .35*(width - (width/2 + key_width/2))), desc(x)),
  arrange(filter(three_ext, x >= (width/2 + key_width/2) + .25*(width - (width/2 + key_width/2))), x),
  c(width - three_point_side_offset, three_point_side_height),
  c(width, three_point_side_height)
)

three_right <- st_polygon(list(as.matrix(three_right)))


deep_three <- rbind(
  arrange(deep_three_arc, x),
  c(width, height),
  c(0, height),
  arrange(deep_three_arc, x)[1,]
)

deep_three <- st_polygon(list(as.matrix(deep_three)))



basic_polys <- st_sf(
  description = c(
    "Left 3", 
    "Middle 3",
    "Right 3",
    "Deep 3",
    "Left Corner 3", 
    "Right Corner 3",
    "Mid-Range Left", 
    "Mid-Range Right",
    "Restricted Area", 
    "In The Paint (Non-RA)",
    "Right Block",
    "Left Block",
    "Mid-Range Center"
  ), 
  geom = c(
    st_geometry(three_left),
    st_geometry(three_middle),
    st_geometry(three_right),
    st_geometry(deep_three),
    st_geometry(left_corner_three),
    st_geometry(right_corner_three),
    st_geometry(mid_range_left),
    st_geometry(mid_range_right),
    st_geometry(restricted_area),
    st_geometry(paint),
    st_geometry(right_block),
    st_geometry(left_block),
    st_geometry(mid_range_center)
  )
) %>%
  transmute(
    shot_zone_basic = description,
    geom
  )

plot_court() +
  geom_sf(data = basic_polys,  
          aes(fill = shot_zone_basic),
          alpha = 0.2,
          show.legend = F) 

```

Lets take the shot data from uconn's games and see how they shot in each zone
```{r}
uconn_data |> 
  select(game_id, scoring_play, score_value, coordinate_x, coordinate_y) -> uconn_data



# adjust shot data to fit the correct coord system
uconn_shot_data <- uconn_data |> 
  mutate(coordinate_x = -1 * abs(coordinate_x) + 47,
         coordinate_y = coordinate_y + 25)

uconn_shots <- st_as_sf(filter(uconn_shot_data, !is.na(coordinate_x)), coords = c("coordinate_y", "coordinate_x")) 




st_join(x = uconn_shots,
        y = basic_polys) |> filter(!is.na(shot_zone_basic)) |> tibble() |> 
  group_by(
    shot_zone_basic
  ) |> 
  summarize(
    num_shots = n(),
    num_makes = sum(scoring_play),
  ) |> 
  mutate(
    percentage = num_makes / num_shots
  ) |> 
  ungroup() -> zones_df


left_join(zones_df,
          basic_polys,
          by = "shot_zone_basic") -> zone_data

zone_data
```

``` {r}
plot_court() +
  geom_sf(data = st_sf(zone_data),  
          aes(fill = percentage),
          color = "white",
          linewidth = .6) + 
  labs(title = "2023 UConn Shooting by Zone (18 game sample size)") + 
  scale_fill_viridis_c() + 
  geom_sf_text(
    data = st_sf(zone_data),
    mapping = aes(label = round(percentage, digits = 2))
  )
```


## Putting it all together

```{r}
# Looking at 2023 shot chart data for our very own UVM catamounts

# filtering out UVM games with no shot location data
ncaa_pbp |> 
  filter(away_team_abbrev == "UVM" | home_team_abbrev == "UVM") |> 
  filter(
    type_text %in% shot_types,
    !is.na(coordinate_x)
  ) |> 
  select(game_id, scoring_play, score_value, coordinate_x, coordinate_y) |> 
  mutate(coordinate_x = -1 * abs(coordinate_x) + 47,
         coordinate_y = coordinate_y + 25) -> UVM_shots 
# 216 shots to work with


# plot the misses and makes on the court
plot_court() +
  geom_point(
    data = UVM_shots,
    mapping = aes(
      color = factor(score_value), 
      x = coordinate_y,
      y = coordinate_x, 
      shape = scoring_play
    ),
    size = 2
  ) + 
  scale_shape_manual(
    values = c(4, 16),
    labels = c("Miss", "Make")
  ) + 
  theme(legend.position = "right") + 
  labs(title = "2023 Vermont Catamounts Shot Chart",
       shape = NULL,
       color = NULL)

# see how they did in each zone

# get data into the right format for the graph
UVM_shots_sf <- st_as_sf(UVM_shots, coords = c("coordinate_y", "coordinate_x")) 


st_join(x = UVM_shots_sf,
        y = basic_polys) |> filter(!is.na(shot_zone_basic)) |> 
  tibble() |> 
  group_by(
    shot_zone_basic
  ) |> 
  summarize(
    num_shots = n(),
    num_makes = sum(scoring_play),
  ) |> 
  mutate(
    percentage = num_makes / num_shots
  ) |> 
  ungroup() -> uvm_zones_df


left_join(uvm_zones_df,
          basic_polys,
          by = "shot_zone_basic") -> uvm_zones_polygons



plot_court() +
  geom_sf(data = st_sf(uvm_zones_polygons),  
          aes(fill = percentage),
          color = "white",
          linewidth = .6) + 
  labs(title = "2023 UVM Shooting by Zone") + 
  scale_fill_viridis_c() + 
  geom_sf_text(
    data = st_sf(uvm_zones_polygons),
    mapping = aes(label = round(percentage, digits = 2))
  )
```
Try it yourself!!



```{r}
# change this team name to be any school you're curious about!
team_name = ""

ncaa_pbp |> 
  filter(away_team_name == team_name | home_team_name == team_name) |> 
  filter(
    type_text %in% shot_types,
    !is.na(coordinate_x)
  ) |> 
  select(game_id, scoring_play, score_value, coordinate_x, coordinate_y) |> 
  mutate(coordinate_x = -1 * abs(coordinate_x) + 47,
         coordinate_y = coordinate_y + 25) -> custom_shots 
# 216 shots to work with


# plot the misses and makes on the court
plot_court() +
  geom_point(
    data = custom_shots,
    mapping = aes(
      color = factor(score_value), 
      x = coordinate_y,
      y = coordinate_x, 
      shape = scoring_play
    ),
    size = 2
  ) + 
  scale_shape_manual(
    values = c(4, 16),
    labels = c("Miss", "Make")
  ) + 
  theme(legend.position = "right") + 
  labs(title = paste("2023", team_name, "Shot Chart"),
       shape = NULL,
       color = NULL)

# see how they did in each zone

# get data into the right format for the graph
custom_shots_sf <- st_as_sf(custom_shots, coords = c("coordinate_y", "coordinate_x")) 


st_join(x = custom_shots_sf,
        y = basic_polys) |> filter(!is.na(shot_zone_basic)) |> 
  tibble() |> 
  group_by(
    shot_zone_basic
  ) |> 
  summarize(
    num_shots = n(),
    num_makes = sum(scoring_play),
  ) |> 
  mutate(
    percentage = num_makes / num_shots
  ) |> 
  ungroup() -> custom_zones_df


left_join(custom_zones_df,
          basic_polys,
          by = "shot_zone_basic") -> custom_zones_polygons



plot_court() +
  geom_sf(data = st_sf(custom_zones_polygons),  
          aes(fill = percentage),
          color = "white",
          linewidth = .6) + 
  labs(title = paste("2023", team_name, "Shooting by Zone")) + 
  scale_fill_viridis_c() + 
  geom_sf_text(
    data = st_sf(custom_zones_polygons),
    mapping = aes(label = round(percentage, digits = 2))
  )


```